Rice yield improvements by plant breeding are offset by decreases over time

### 2017-06-23 output/breed_paper.Rnw
Adjusts figure 2 to be panel plot

<<fig=TRUE, echo=FALSE>>=
yrs <- 1975:2025
pp <- posterior_linpred(RES_full, re.form = NA,
                        newdata = data.frame(release_yr_c = yrs - 2000,
                                             yrs_in_trial = 0),
                        prob = 0.5)
pp <-  (pp * scl + center) / 1000
ints = apply(pp, 2, quantile, p = c(0.05, 0.95))

@@

plot(yield_kg ~ release_yr, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Release year",
     ylab = expression("Grain yield ( Mg ha"^-1~")"),
     ylim = c(8, 12.5),
     xlim = c(1980,2019))
abline(a = ((((fixef(RES_full)[1] * scl) + center) ) -
           (2000 * release_yr_coef)) / 1000,
       b = release_yr_coef/1000)
abline(lm(yield_kg/1000 ~ release_yr, data = mod_data[idx,]), lty = 2)
with(plot_data, segments(x0 = release_yr, y0=yield_kg + yield_se,
                         y1 = yield_kg - yield_se))
text(x = plot_data$release_yr-0.4, y = plot_data$yield_kg + 0.07,
     labels = gsub("M", "M-", row.names(plot_data)),
     pos = 4, cex = 0.8)

# legend("bottomright", legend = c("w/ years since release adj.",
#                                  "w/o years since release adj.",
#                                  "90% Credible interval"),
#        lty = c(1,2, NA),
#        bty = "n",
#        fill = c(NA,NA, "grey"),
#        border = NA,
#        merge = TRUE,
#        cex = 0.8)

legend("bottomright", legend = c("w/ years since release adj.",
                                 "w/o years since release adj."),
       lty = c(1,2),
       bty = "n",
       border = NA,
       merge = TRUE,
       cex = 0.8)
# polygon(x = c(yrs, rev(yrs)),
        # y = c(ints[1,],rev(ints[2,])),
        # border = NA,
        # col = rgb(0,0,0,0.1))

# text(1990, 8.0, bquote(.(round(release_yr_coef))~kg~ha^-1~year^-1))
@
\caption{The average yield ($\bullet$) and standard error (|) of
medium grain rice cultivars by year of release. For display purposes,
only data for experimental plots planted at the California Cooperative
Rice Research Foundation Rice Experiment Station (California, USA)
over the years \Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)} are
included. Ignoring difference in years since release (dashed line)
suggests an increase in yield potential over time. However, correcting
the comparison to account for yield erosion over time (solid line)
suggests the apparent increase might be due to differences in years
since release.}
\label{fig:release}
\end{figure}

@@

@
\caption{The relationship between yield and the years since release
  for medium grain rice varieties planted at the Rice Experiment Station
  (RES), California, USA. The estimated erosion of yield over time
  (averaged over differences in cultivar and year) is
  estimated to be \Sexpr{yrs_coef} kg ha\tsups{-1}
  year\tsups{-1}. Points ($bullet$) and bars (|) represent the mean and standard
  error of the raw data, respectively. The
  shaded area is the 90\% credible interval.}
\label{fig:RES_year}
\end{figure}

>>>>

<<fig=TRUE, echo=FALSE, width = 9, height = 6>>=
yrs <- 1975:2025

pp <- posterior_linpred(RES_rel_fit, re.form = NA,
                        newdata = data.frame(release_yr_c = yrs - 2000,
                                             yrs_in_trial = 0),
                        prob = 0.5)
pp <-  (pp * scl + center) / 1000
ints1 = apply(pp, 2, quantile, p = c(0.05, 0.95))

pp <- posterior_linpred(RES_full, re.form = NA,
                        newdata = data.frame(release_yr_c = yrs - 2000,
                                             yrs_in_trial = 0),
                        prob = 0.5)
pp <-  (pp * scl + center) / 1000
ints2 = apply(pp, 2, quantile, p = c(0.05, 0.95))

@@

# get corrected yield
wo_erosion = expand.grid(id = row.names(plot_data),
            year = comp_yrs,
            release_yr_c = 0,
            yrs_in_trial = 0)
wo_erosion$corr_yield = colMeans(posterior_predict(RES_full,
                               newdata = wo_erosion)) + center/1000
plot_data2 = prep_plot_data(wo_erosion$corr_yield,
                            wo_erosion$id)
plot_data2$release_yr = plot_data$release_yr

par(mfrow = c(1,2),
    mar = c(4,0,2,0.75),
    oma = c(4,4,0.5,0.5))

plot(yield_kg ~ release_yr, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Release year",
     ylab = "", 
     ylim = c(8, 12.5),
     xaxt = "n",
     xlim = c(1980,2020))
axis(1, at = seq(1980, 2020, by = 10))
mtext(text = expression("Grain yield ( Mg ha"^-1~")"), side = 2, line = 2.5)
abline(a = ((fixef(RES_rel_fit)[1] * scl) + center) / 1000 -
      (2000 * fixef(RES_rel_fit)[2]),
       b = fixef(RES_rel_fit)[2],
       lty = 2)
polygon(x = c(yrs, rev(yrs)),
        y = c(ints1[1,],rev(ints1[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

with(plot_data, segments(x0 = release_yr, y0=yield_kg + yield_se,
                         y1 = yield_kg - yield_se))
text(x = plot_data$release_yr-0.4, y = plot_data$yield_kg + 0.07,
     labels = gsub("M", "M-", row.names(plot_data)),
     pos = 4, cex = 0.75)
text(x = 2000, y = 8, bquote(.(round(fixef(RES_rel_fit)[2] * scl, 1))~kg~ha^-1~yr^-1))
mtext("a.", adj = 0)

plot(mean ~ release_yr, data = plot_data2, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Release year",
     ylab = "", #expression("Grain yield -corrected( Mg ha"^-1~")"),
     yaxt = "n",
     ylim = c(8, 12.5),
     xaxt = "n",
     xlim = c(1980,2020))
axis(1, at = seq(1980, 2020, by = 10))
abline(a = ((((fixef(RES_full)[1] * scl) + center) ) -
           (2000 * release_yr_coef)) / 1000,
       b = release_yr_coef/1000,
       lty = 2)
with(plot_data2, segments(x0 = release_yr, y0 = mean + se,
                         y1 = mean - se))
text(x = plot_data2$release_yr-0.4, y = plot_data2$mean + 0.07,
     labels = gsub("M", "M-", row.names(plot_data2)),
     pos = 4, cex = 0.75)
text(x = 2000, y = 8, bquote(.(release_yr_coef)~kg~ha^-1~yr^-1))
mtext("b.", adj = 0)
polygon(x = c(yrs, rev(yrs)),
        y = c(ints2[1,],rev(ints2[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

@
\caption{The average yield ($\bullet$) and standard error (|) of
medium grain rice cultivars by year of release (a.) without correcting
for yield erosion over time and (b.) accounting for yield erosion over
time. The trend without correcting for yield erosion (a.) was estimated by
omitting the year of each cultivar's release from the
model. Yield erosion was corrected for by including the years since
release (b.) in the model. predictions assuming no yield erosion
were generated from the model which included both the year of release
and the years since release. For display purposes, only data for
experimental plots planted at the California Cooperative Rice Research
Foundation Rice Experiment Station (California, USA) over the years
\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)} are included.}
\label{fig:release}
\end{figure}

@@

@
\caption{The relationship between yield and the years since release
  for medium grain rice varieties planted at the Rice Experiment Station
  (RES), California, USA. The estimated erosion of yield over time
  (averaged over differences in cultivar and year) is
  estimated to be \Sexpr{yrs_coef} kg ha\tsups{-1}
  year\tsups{-1}. Points ($\bullet$) and bars (|) represent the mean and standard
  error of the raw data, respectively. The
  shaded area is the 90\% credible interval.}
\label{fig:RES_year}
\end{figure}
